---
title: "Conversion long/wide format"
format: html
author: "baptiste"
date: today
---

[R utility functions](_fun.R) used below.

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(grid)
library(patchwork)
read_chunk('_fun.R')
source("_fun.R")
```


## Reading long-format T-matrix files

Some programs output T-matrices in long format, similar to the example below:

```{r}
d <- read.table('data/tmat_Au20x40_Nmax3.tmat')
names(d) <- c('s','sp','l','lp','m','mp','Tr','Ti')
head(d)
```

Note that one advantage of this format is that only non-zero entries need to be provided, similar to a sparse matrix definition. This is useful for EBCM-like analytical methods where particle symmetries dictate that some T-matrix elements be identically zero (plane of symmetry, axis of symmetry, etc.).

With such data, it can be useful to add combined indices $p$ (blocks) and $q$ (full matrix), as illustrated below:

```{r}
d <- tmatrix_combinedindex(d)
head(d)
```

## Reading wide data and converting to long format

The function `read_treams()` reads a standard HDF5-stored T-matrix and returns it in long format, including the computation's `method_parameters` fields.

```{r}
d <- read_treams('data/SPH-DE~4.H5')
head(d)
```


## Convert to wide

If we wish to convert to a true matrix representation, the following function adds all the missing zeros and places the entries at the [combined indices](indexing.html) $(q,q')$,

```{r wide, echo=1}
w <- tmatrix_wide(d)

prettymat(w, coldots=3)
```


